An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-11 EconGrowth-R-Functions.Rnw
Re-arranged the CES parameter tables in the appendix of the paper.

############################
# Aggregates the CES results for gamma, delta_1, and delta into a table for the given energyType.
##
cesParamsTableA <- function(energyType){
  dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
  colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
                         " ", "$\\lambda$",   " ",
                         " ", "$\\delta_1$",  " ",
                         " ", "$\\delta$",    " ",
                         " ", "$\\sigma_1$",  " ",
                         " ", "$\\sigma$",    " ")
  rownames(dataCES) <- countryAbbrevs
  tableCESa <- xtable(dataCES[,c(1,2,3, 7,8,9, 10,11,12)], #Picks up gamma, delta_1, and delta
                      caption=paste("CES model (with $", tolower(energyType), "$). $\\gamma$, $\\delta_1$, and $\\delta$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                      label=paste("tab:CES_Parameters_With_", energyType, sep=""),
                      digits = c(1, 2,2,2, 2,2,2, 2,2,2),
                      align = "r|ccc|ccc|ccc" #Sets alignment of the numbers in the columns
                      ) 
  return(tableCESa)
}

############################
# Aggregates the CES results for lambda, sigma_1, and sigma into a table for the given energyType.
##
cesParamsTableB <- function(energyType){
  dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
  colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
                         " ", "$\\lambda$",   " ",
                         " ", "$\\delta_1$",  " ",
                         " ", "$\\delta$",    " ",
                         " ", "$\\sigma_1$",  " ",
                         " ", "$\\sigma$",    " ")
  rownames(dataCES) <- countryAbbrevs
  tableCESb <- xtable(dataCES[,c(4,5,6, 13,14,15, 16,17,18)], #Picks up lambda, sigma_1, and sigma
                      caption=paste("CES model (with $", tolower(energyType), "$). $\\lambda$, $\\sigma_1$, and $\\sigma$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                      label=paste("tab:CES_Parameters_With_", energyType, sep=""),
                      digits = c(1, 4,4,4, 2,2,2, 2,2,2),
                      align = "r|ccc|ccc|ccc"  #Sets alignment of the numbers in the columns
                      )
  return(tableCESb)
}

@@

<<AIC_Table_Functions, eval=TRUE>>=
###############################
# Creates an xtable object that holds the AIC values for each parameter estimation that we include.
##
createAICTable <- function(){
  ######################
  # Single-factor models
  ######################
  # Single-factor with K
  sfKModels <- lapply(countryAbbrevs, singleFactorModel, factor="K")
  aicSFk <- data.frame(lapply(sfKModels, AIC))
  rownames(aicSFk) <- "SF$k$"
  # Single-factor with L
  sfLModels <- lapply(countryAbbrevs, singleFactorModel, factor="L")
  aicSFl <- data.frame(lapply(sfLModels, AIC))
  rownames(aicSFl) <- "SF$l$"
  # Single-factor with Q
  sfQModels <- lapply(countryAbbrevs, singleFactorModel, factor="Q")
  aicSFq <- data.frame(lapply(sfQModels, AIC))
  rownames(aicSFq) <- "SF$q$"
  # Single-factor with X
  sfXModels <- lapply(countryAbbrevs, singleFactorModel, factor="X")
  aicSFx <- data.frame(lapply(sfXModels, AIC))
  rownames(aicSFx) <- "SF$x$"
  # Single-factor with U
  aicSFu <- cbind(US=AIC(singleFactorModel("US", "U")), 
                  UK=AIC(singleFactorModel("UK", "U")), 
                  JP=AIC(singleFactorModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicSFu) <- "SF$u$"
  ######################
  # Cobb-Douglas models
  ######################
  # Cobb-Douglas without energy
  cdModels <- lapply(countryAbbrevs, cobbDouglasModel)
  aicCD <- data.frame(lapply(cdModels, AIC))
  rownames(aicCD) <- "CD"
  # Cobb-Douglas with Q
  cdQModels <- lapply(countryAbbrevs, cobbDouglasModel, energyType="Q")
  aicCDq <- data.frame(lapply(cdQModels, AIC))
  rownames(aicCDq) <- "CD$q$"
  # Cobb-Douglas with X
  cdXModels <- lapply(countryAbbrevs, cobbDouglasModel, energyType="X")
  aicCDx <- data.frame(lapply(cdXModels, AIC))
  rownames(aicCDx) <- "CD$x$"
  # Cobb-Douglas with U
  aicCDu <- cbind(US=AIC(cobbDouglasModel("US", "U")), 
                  UK=AIC(cobbDouglasModel("UK", "U")), 
                  JP=AIC(cobbDouglasModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicCDu) <- "CD$u$"
  ######################
  # CES models
  ######################

>>>>

# ############################
# # Aggregates the CES results for gamma, delta_1, and delta into a table for the given energyType.
# ##
# cesParamsTableA <- function(energyType){
#   dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
#   colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
#                          " ", "$\\lambda$",   " ",
#                          " ", "$\\delta_1$",  " ",
#                          " ", "$\\delta$",    " ",
#                          " ", "$\\sigma_1$",  " ",
#                          " ", "$\\sigma$",    " ")
#   rownames(dataCES) <- countryAbbrevs
#   tableCESa <- xtable(dataCES[,c(1,2,3, 7,8,9, 10,11,12)], #Picks up gamma, delta_1, and delta
#                       caption=paste("CES model (with $", tolower(energyType), "$). $\\gamma$, $\\delta_1$, and $\\delta$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
#                       label=paste("tab:CES_Parameters_With_", energyType, sep=""),
#                       digits = c(1, 2,2,2, 2,2,2, 2,2,2),
#                       align = "r|ccc|ccc|ccc" #Sets alignment of the numbers in the columns
#                       ) 
#   return(tableCESa)
# }
# 
# ############################
# # Aggregates the CES results for lambda, sigma_1, and sigma into a table for the given energyType.
# ##
# cesParamsTableB <- function(energyType){
#   dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
#   colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
#                          " ", "$\\lambda$",   " ",
#                          " ", "$\\delta_1$",  " ",
#                          " ", "$\\delta$",    " ",
#                          " ", "$\\sigma_1$",  " ",
#                          " ", "$\\sigma$",    " ")
#   rownames(dataCES) <- countryAbbrevs
#   tableCESb <- xtable(dataCES[,c(4,5,6, 13,14,15, 16,17,18)], #Picks up lambda, sigma_1, and sigma
#                       caption=paste("CES model (with $", tolower(energyType), "$). $\\lambda$, $\\sigma_1$, and $\\sigma$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
#                       label=paste("tab:CES_Parameters_With_", energyType, sep=""),
#                       digits = c(1, 4,4,4, 2,2,2, 2,2,2),
#                       align = "r|ccc|ccc|ccc"  #Sets alignment of the numbers in the columns
#                       )
#   return(tableCESb)
# }
#

############################
# Aggregates the CES results for gamma, delta_1, and delta into a table for the given energyType.
##
cesParamsTableA <- function(energyType){
  dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
  colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
                         " ", "$\\lambda$",   " ",
                         " ", "$\\delta_1$",  " ",
                         " ", "$\\delta$",    " ",
                         " ", "$\\sigma_1$",  " ",
                         " ", "$\\sigma$",    " ")
  rownames(dataCES) <- countryAbbrevs
  tableCESa <- xtable(dataCES[,c(4,5,6, 10,11,12, 16,17,18)], #Picks up lambda, delta, sigma
                      caption=paste("CES model (with $", tolower(energyType), "$). $\\lambda$, $\\delta$, and $\\sigma$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                      label=paste("tab:CES_Parameters_A_With_", energyType, sep=""),
                      digits = c(1, 4,4,4, 2,2,2, 2,2,2),
                      align = "r|ccc|ccc|ccc" #Sets alignment of the numbers in the columns
                      ) 
  return(tableCESa)
}

############################
# Aggregates the CES results for lambda, sigma_1, and sigma into a table for the given energyType.
##
cesParamsTableB <- function(energyType){
  dataCES <- do.call("rbind", lapply(countryAbbrevs, cesCountryRow, energyType=energyType))
  colnames(dataCES) <- c(" ", "$\\gamma$",    " ", 
                         " ", "$\\lambda$",   " ",
                         " ", "$\\delta_1$",  " ",
                         " ", "$\\delta$",    " ",
                         " ", "$\\sigma_1$",  " ",
                         " ", "$\\sigma$",    " ")
  rownames(dataCES) <- countryAbbrevs
  tableCESb <- xtable(dataCES[,c(1,2,3, 7,8,9, 13,14,15)], #Picks up gamma, delta_1, and sigma_1
                      caption=paste("CES model (with $", tolower(energyType), "$). $\\gamma$, $\\delta_1$, and $\\sigma_1$ parameters for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                      label=paste("tab:CES_Parameters_B_With_", energyType, sep=""),
                      digits = c(1, 2,2,2, 2,2,2, 2,2,2),
                      align = "r|ccc|ccc|ccc"  #Sets alignment of the numbers in the columns
                      )
  return(tableCESb)
}

@@

<<AIC_Table_Functions, eval=TRUE>>=
###############################
# Creates an xtable object that holds the AIC values for each parameter estimation that we include.
##
createAICTable <- function(){
  ######################
  # Single-factor models
  ######################
  # Single-factor with K
  sfKModels <- lapply(countryAbbrevs, singleFactorModel, factor="K")
  aicSFk <- data.frame(lapply(sfKModels, AIC))
  rownames(aicSFk) <- "SF$k$"
  # Single-factor with L
  sfLModels <- lapply(countryAbbrevs, singleFactorModel, factor="L")
  aicSFl <- data.frame(lapply(sfLModels, AIC))
  rownames(aicSFl) <- "SF$l$"
  # Single-factor with Q
  sfQModels <- lapply(countryAbbrevs, singleFactorModel, factor="Q")
  aicSFq <- data.frame(lapply(sfQModels, AIC))
  rownames(aicSFq) <- "SF$q$"
  # Single-factor with X
  sfXModels <- lapply(countryAbbrevs, singleFactorModel, factor="X")
  aicSFx <- data.frame(lapply(sfXModels, AIC))
  rownames(aicSFx) <- "SF$x$"
  # Single-factor with U
  aicSFu <- cbind(US=AIC(singleFactorModel("US", "U")), 
                  UK=AIC(singleFactorModel("UK", "U")), 
                  JP=AIC(singleFactorModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicSFu) <- "SF$u$"
  ######################
  # Cobb-Douglas models
  ######################
  # Cobb-Douglas without energy
  cdModels <- lapply(countryAbbrevs, cobbDouglasModel)
  aicCD <- data.frame(lapply(cdModels, AIC))
  rownames(aicCD) <- "CD"
  # Cobb-Douglas with Q
  cdQModels <- lapply(countryAbbrevs, cobbDouglasModel, energyType="Q")
  aicCDq <- data.frame(lapply(cdQModels, AIC))
  rownames(aicCDq) <- "CD$q$"
  # Cobb-Douglas with X
  cdXModels <- lapply(countryAbbrevs, cobbDouglasModel, energyType="X")
  aicCDx <- data.frame(lapply(cdXModels, AIC))
  rownames(aicCDx) <- "CD$x$"
  # Cobb-Douglas with U
  aicCDu <- cbind(US=AIC(cobbDouglasModel("US", "U")), 
                  UK=AIC(cobbDouglasModel("UK", "U")), 
                  JP=AIC(cobbDouglasModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicCDu) <- "CD$u$"
  ######################
  # CES models
  ######################
  cesQModels <- lapply(countryAbbrevs, cesModel, energyType="Q")
  aicCESq <- data.frame(lapply(cesQModels, AIC))
  rownames(aicCESq)

### 2013-04-11 EconGrowth.Rnw
Re-arranged the CES parameter tables in the appendix of the paper.

\section*{To-do List}
\begin{itemize}
\item Reese: Add text from Word verson of paper to the LaTeX version.
\item Heun: Add CES production function fits
      \begin{itemize}
      \item \st{GDP comparison graph}
      \item \st{Convert $\rho$ variables in the CES production function to $\sigma$}
      \item Write up the decisions about parameter estimation for the CES production function fits
      \item \st{Create parameter graphs for each country}
      \item \st{Parameter tables for each energy type in the appendix}
      \item Fix the problems with $\sigma_1$ and $\sigma$ values in the CES parameter tables in the appendix
      \item AIC rows in the AIC table
      \end{itemize}
\item Heun: Deal with missing covariance numbers when columns have unequal length. Use the \texttt{subset} command.
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the Cobb-Douglas with energy rows. Asked Pruim about this via email on 4 Feb 2013 but have not heard a response.
      \begin{itemize}
      \item ZA: lower bound on $\lambda$ and upper bound on $\alpha$
      \item ZM: lower bound on $\lambda$
      \end{itemize}
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the LINEX model
      \begin{itemize}
      \item Zambia, upper bound on $c_t$
      \end{itemize}
\item Heun: Fix warnings of the form ``Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562'' in the code that generates the Cobb-Douglas with energy fits.
\item Heun: x-axis labels only on bottom for 9-panel lattice graphs
\item Heun: adjust y axis limits on $a_0$ and $c_t$ parameter graphs for the LINEX model AFTER we finish debugging the LINEX models.
\item de Wit: provide funding acknowledgement, if necessary.
\item de Wit: review, critique, and contextualize.
\item \st{Heun: Ask Martin if he can supply an ``Economics'' affiliation. Will probably make a stronger paper.}
\item \st{Heun: Create a table of AIC values for each fit.}
\item \st{Heun: Add $u$ predictions for Cobb-Douglas}
\item \st{Heun: Move tables to an appendix at the end.}
\item \st{Heun: Add acknowledgements (Dad, Pruim)}
\item \st{Heun: Add $u$ parameter table and graph}
\item \st{Heun: Add covariance metrics.}
\item \st{Heun: Add single factor fits}
      \begin{itemize}
      \item \st{Correct sizes for parameter graphs for each country}
      \item \st{Parameter graphs ($\lambda$ and $m$) for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{GDP comparison graph}
      \item \st{AIC rows in the AIC table}
      \end{itemize}
\item \st{Heun: Add setup variables for figure height and width where not already included.}
\item \st{Heun: Add LINEX production function fits}
      \begin{itemize}
      \item \st{Parameter graphs for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{AIC rows in the AIC table}
      \item \st{GDP comparison graph}
      \end{itemize}
\end{itemize}

@@

<<test3dPlot, eval=TRUE>>=
sf3DSSEGraph("US", "Q")
@

@@

<<CES_Params_Table, results='asis'>>=
#printCESParamsTable("Q"); printCESParamsTable("X"); printCESParamsTable("U")
printCESParamsTableA("Q"); printCESParamsTableB("Q")
printCESParamsTableA("X"); printCESParamsTableB("X")
printCESParamsTableA("U"); printCESParamsTableB("U")
@

>>>>

\section*{To-do List}
\begin{itemize}
\item Reese: Add text from Word verson of paper to the LaTeX version.
\item Heun: Add CES production function fits
      \begin{itemize}
      \item \st{GDP comparison graph}
      \item \st{Convert $\rho$ variables in the CES production function to $\sigma$}
      \item Write up the decisions about parameter estimation for the CES production function fits
      \item \st{Create parameter graphs for each country}
      \item \st{Parameter tables for each energy type in the appendix}
      \item Fix the problems with $\sigma_1$ and $\sigma$ values in the CES parameter tables in the appendix
      \item AIC rows in the AIC table
      \item \st{Re-organize CES tables into (a) $\lambda$, $\delta$, $\sigma$ and (b) $\gamma$, $\delta_1$, $\sigma_1$}
      \end{itemize}
\item Heun: Deal with missing covariance numbers when columns have unequal length. Use the \texttt{subset} command?
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the Cobb-Douglas with energy rows. Asked Pruim about this via email on 4 Feb 2013 but have not heard a response.
      \begin{itemize}
      \item ZA: lower bound on $\lambda$ and upper bound on $\alpha$
      \item ZM: lower bound on $\lambda$
      \end{itemize}
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the LINEX model
      \begin{itemize}
      \item Zambia, upper bound on $c_t$
      \end{itemize}
\item Heun: Fix warnings of the form ``Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562'' in the code that generates the Cobb-Douglas with energy fits.
\item Heun: x-axis labels only on bottom for 9-panel lattice graphs
\item Heun: adjust y axis limits on $a_0$ and $c_t$ parameter graphs for the LINEX model AFTER we finish debugging the LINEX models.
\item de Wit: provide funding acknowledgement, if necessary.
\item de Wit: review, critique, and contextualize.
\item \st{Heun: Ask Martin if he can supply an ``Economics'' affiliation. Will probably make a stronger paper.}
\item \st{Heun: Create a table of AIC values for each fit.}
\item \st{Heun: Add $u$ predictions for Cobb-Douglas}
\item \st{Heun: Move tables to an appendix at the end.}
\item \st{Heun: Add acknowledgements (Dad, Pruim)}
\item \st{Heun: Add $u$ parameter table and graph}
\item \st{Heun: Add covariance metrics.}
\item \st{Heun: Add single factor fits}
      \begin{itemize}
      \item \st{Correct sizes for parameter graphs for each country}
      \item \st{Parameter graphs ($\lambda$ and $m$) for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{GDP comparison graph}
      \item \st{AIC rows in the AIC table}
      \end{itemize}
\item \st{Heun: Add setup variables for figure height and width where not already included.}
\item \st{Heun: Add LINEX production function fits}
      \begin{itemize}
      \item \st{Parameter graphs for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{AIC rows in the AIC table}
      \item \st{GDP comparison graph}
      \end{itemize}
\end{itemize}

@@

@@

<<CES_Params_Table, results='asis'>>=
printCESParamsTableA("Q"); printCESParamsTableB("Q")
printCESParamsTableA("X"); printCESParamsTableB("X")
printCESParamsTableA("U"); printCESParamsTableB("U")
@
